---
title: "Homework 02"
author: "yourname"
date: "Septemeber 16, 2018"
output:
  pdf_document: default
---

\newcommand{\mat}[1]{\boldsymbol{#1}} 
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\rv}[1]{\underline{#1}}

```{r setup, include=FALSE}
pacman::p_load("arm",
               "data.table",
               "Cairo",
               "faraway",
               "foreign",
               "tidyverse",
               "knitr",
               'VIM',
               "kableExtra")
opts_chunk$set(echo = F,dev="CairoPNG",fig.align = "center", 
                      fig.width = 5.656, fig.height = 4, global.par = TRUE,
               warning = FALSE, message=FALSE)
par (mar=c(3,3,2,1), mgp=c(2,.7,0), tck=-.01)
```

# Introduction 
In homework 2 you will fit many regression models.  You are welcome to explore beyond what the question is asking you.  

Please come see us we are here to help.

## Data analysis 

### Analysis of earnings and height data

The folder `earnings` has data from the Work, Family, and Well-Being Survey (Ross, 1990).
You can find the codebook at http://www.stat.columbia.edu/~gelman/arm/examples/earnings/wfwcodebook.txt
```{r}
gelman_dir <- "http://www.stat.columbia.edu/~gelman/arm/examples/"
heights    <- read.dta (paste0(gelman_dir,"earnings/heights.dta"),warn.missing.labels = T)
```

Pull out the data on earnings, sex, height, and weight.
1. In R, check the dataset and clean any unusually coded data.
```{r echo=FALSE, fig.height=3.3,dpi=200}
info=NULL
for(i in 1:length(names(heights))){
  num_of_uniquevalue=length(unique(heights[,i]))
  Na = sum(is.na(heights[,i]))
  Ze = sum(heights[,i]==0,na.rm =T)
  P_na = (Na+Ze)/length(heights[,i])
  info = cbind(info,matrix(c(num_of_uniquevalue,Na,Ze,P_na),ncol = 1))
}
rownames(info)=c('Number of Unique Value','Number of Missing Value','Number of 0','% of Missing Value')
colnames(info)=names(heights)
kable(info,caption = "Quality of Data",align = "c",digits = 2,
      format = 'latex',booktabs=T,longtable=T)
matrixplot(heights)
```
\qquad Above chart and figure show the distribution of missing value amoung variables, in the figure, the red means there is a missing value, and for the rest, the darker the color, the higher the value.  

\qquad So according to the analysis above, the most of the missing value ,which is represented by 'NA' and '0', contained in variable "earning", about $41\%$ of data doesn`t have variable 'earning', which is reasonable, beacause earning is kond of a private quetion, due to the extent of missing, we can hardly apply any imputaion, so simply delete it.  
```{r fig.height=3.3,dpi=200}
for(i in 1:length(names(heights))){
  heights[which(heights[,i]==0),i]=NaN
}
hi = na.omit(heights)
info=NULL
for(i in 1:length(names(hi))){
  num_of_uniquevalue=length(unique(hi[,i]))
  Na = sum(is.na(hi[,i]))
  Ze = sum(hi[,i]==0)
  P_na = (Na+Ze)/length(hi[,i])
  info = cbind(info,matrix(c(num_of_uniquevalue,Na,Ze,P_na),ncol = 1))
}
rownames(info)=c('Number of Unique Value','Number of Missing Value','Number of 0','% of Missing Value')
colnames(info)=names(hi)
kable(info,caption = "Quality of Data",digits = 2,
      align = "c",format = 'latex',booktabs=T,longtable = T)
matrixplot(hi)
```
\qquad So now the data is clean and ready to be analysisd.  
2. Fit a linear regression model predicting earnings from height. What transformation should you perform in order to interpret the intercept from this model
as average earnings for people with average height?  

\qquad Look at the data, we can find three variables that related to height, "height 1", "height 2" and "height", after reading the original article, we know that real height should be "hight 1" feet and "height 2" inches, so $Real Height=12*height_1+height_2$, which is exactly the variable "height".  
```{r}
hi[,'h_over_m']=hi$height - mean(hi$height)
ggplot(hi,aes(y=earn,x = h_over_m)) +
  geom_point() + geom_smooth(se=F)
```  
\qquad Look at the plot of height over average against earning, we can easily find some outliers, and a linear relationship may seams wield, so we apply some non-linear transformation, first we consider use $\log(earn)$ replace earn to eliminate the effect of outliers.  As for the predictor, we try to include the polynomial in our model.  
```{r fig.width= 8}
pnl=ggplot(hi,aes(y=earn,x = h_over_m)) +
  geom_point() + geom_smooth(se=F)
pl=ggplot(hi,aes(y=log(earn,10),x = h_over_m)) +
  geom_point() + geom_smooth(se=F)
gridExtra::grid.arrange(pnl,pl,ncol=2)
```
\qquad From above plot, we can tell that replacing earn by $\log_{10}Earn$ does comdence the data and alleviate the effect of outliers a bit. Than apply regression analysis.  
```{r}
mo = lm(log(earn,10)~h_over_m+I(exp(h_over_m)),data = hi)
summary(mo)
ggplot()+geom_point(aes(y=mo$residuals,x=mo$fitted.values))
car::qqPlot(mo$residuals)
```

3. Fit some regression models with the goal of predicting earnings from some
combination of sex, height, and weight. Be sure to try various transformations and interactions that might make sense. Choose your preferred model and justify.

```{r}

```

4. Interpret all model coefficients.

```{r}

```

5. Construct 95% confidence interval for all model coefficients and discuss what they mean.

```{r}

```


### Analysis of mortality rates and various environmental factors

The folder `pollution` contains mortality rates and various environmental factors from 60 U.S. metropolitan areas from McDonald, G.C. and Schwing, R.C. (1973) 'Instabilities of regression estimates relating air pollution to mortality', Technometrics, vol.15, 463-482. 

Variables, in order:

* PREC   Average annual precipitation in inches
* JANT   Average January temperature in degrees F
* JULT   Same for July
* OVR65  % of 1960 SMSA population aged 65 or older
* POPN   Average household size
* EDUC   Median school years completed by those over 22
* HOUS   % of housing units which are sound & with all facilities
* DENS   Population per sq. mile in urbanized areas, 1960
* NONW   % non-white population in urbanized areas, 1960
* WWDRK  % employed in white collar occupations
* POOR   % of families with income < $3000
* HC     Relative hydrocarbon pollution potential
* NOX    Same for nitric oxides
* SO@    Same for sulphur dioxide
* HUMID  Annual average % relative humidity at 1pm
* MORT   Total age-adjusted mortality rate per 100,000

For this exercise we shall model mortality rate given nitric oxides, sulfur dioxide, and hydrocarbons as inputs. This model is an extreme oversimplification as it combines all sources of mortality and does not adjust for crucial factors such as age and smoking. We use it to illustrate log transformations in regression.

```{r}
gelman_dir   <- "http://www.stat.columbia.edu/~gelman/arm/examples/"
pollution    <- read.dta (paste0(gelman_dir,"pollution/pollution.dta"))
```

1. Create a scatterplot of mortality rate versus level of nitric oxides. Do you think linear regression will fit these data well? Fit the regression and evaluate a residual plot from the regression.

```{r}

```

2. Find an appropriate transformation that will result in data more appropriate for linear regression. Fit a regression to the transformed data and evaluate the new residual plot.

```{r}

```

3. Interpret the slope coefficient from the model you chose in 2.

```{r}

```

4. Construct 99% confidence interval for slope coefficient from the model you chose in 2 and interpret them.

```{r}

```

5. Now fit a model predicting mortality rate using levels of nitric oxides, sulfur dioxide, and hydrocarbons as inputs. Use appropriate transformations when
helpful. Plot the fitted regression model and interpret the coefficients.

```{r}

```

6. Cross-validate: fit the model you chose above to the first half of the data and then predict for the second half. (You used all the data to construct the model in 4, so this is not really cross-validation, but it gives a sense of how the steps of cross-validation can be implemented.)

```{r}

```

### Study of teenage gambling in Britain

```{r,message =FALSE}
data(teengamb)
#?teengamb
```

1. Fit a linear regression model with gamble as the response and the other variables as predictors and interpret the coefficients. Make sure you rename and transform the variables to improve the interpretability of your regression model.

```{r}

```

2. Create a 95% confidence interval for each of the estimated coefficients and discuss how you would interpret this uncertainty.

```{r}

```

3. Predict the amount that a male with average status, income and verbal score would gamble along with an appropriate 95% CI.  Repeat the prediction for a male with maximal values of status, income and verbal score.  Which CI is wider and why is this result expected?

```{r}

```

### School expenditure and test scores from USA in 1994-95

```{r}
data(sat)
#?sat
```

1. Fit a model with total sat score as the outcome and expend, ratio and salary as predictors.  Make necessary transformation in order to improve the interpretability of the model.  Interpret each of the coefficient.

```{r}

```

2. Construct 98% CI for each coefficient and discuss what you see.

```{r}

```

3. Now add takers to the model.  Compare the fitted model to the previous model and discuss which of the model seem to explain the outcome better?

```{r}

```

# Conceptual exercises.

### Special-purpose transformations:

For a study of congressional elections, you would like a measure of the relative amount of money raised by each of the two major-party candidates in each district. Suppose that you know the amount of money raised by each candidate; label these dollar values $D_i$ and $R_i$. You would like to combine these into a single variable that can be included as an input variable into a model predicting vote share for the Democrats.

Discuss the advantages and disadvantages of the following measures:

* The simple difference, $D_i-R_i$

* The ratio, $D_i/R_i$

* The difference on the logarithmic scale, $log D_i-log R_i$ 

* The relative proportion, $D_i/(D_i+R_i)$.


### Transformation 

For observed pair of $\mathrm{x}$ and $\mathrm{y}$, we fit a simple regression model 
$$\mathrm{y}=\alpha + \beta \mathrm{x} + \mathrm{\epsilon}$$ 
which results in estimates $\hat{\alpha}=1$, $\hat{\beta}=0.9$, $SE(\hat{\beta})=0.03$, $\hat{\sigma}=2$ and $r=0.3$.

1. Suppose that the explanatory variable values in a regression are transformed according to the $\mathrm{x}^{\star}=\mathrm{x}-10$ and that $\mathrm{y}$ is regressed on $\mathrm{x}^{\star}$.  Without redoing the regression calculation in detail, find $\hat{\alpha}^{\star}$, $\hat{\beta}^{\star}$, $\hat{\sigma}^{\star}$, and $r^{\star}$.  What happens to these quantities when $\mathrm{x}^{\star}=10\mathrm{x}$ ? When $\mathrm{x}^{\star}=10(\mathrm{x}-1)$?


2. Now suppose that the response variable scores are transformed according to the formula
$\mathrm{y}^{\star\star}= \mathrm{y}+10$ and that $\mathrm{y}^{\star\star}$ is regressed on $\mathrm{x}$.  Without redoing the regression calculation in detail, find $\hat{\alpha}^{\star\star}$, $\hat{\beta}^{\star\star}$, $\hat{\sigma}^{\star\star}$, and $r^{\star\star}$.  What happens to these quantities when $\mathrm{y}^{\star\star}=5\mathrm{y}$ ? When $\mathrm{y}^{\star\star}=5(\mathrm{y}+2)$?


3. In general, how are the results of a simple regression analysis affected by linear transformations of $\mathrm{y}$ and $\mathrm{x}$?



4. Suppose that the explanatory variable values in a regression are transformed according to the $\mathrm{x}^{\star}=10(\mathrm{x}-1)$ and that $\mathrm{y}$ is regressed on $\mathrm{x}^{\star}$.  Without redoing the regression calculation in detail, find $SE(\hat{\beta}^{\star})$ and $t^{\star}_0= \hat{\beta}^{\star}/SE(\hat{\beta}^{\star})$.


5. Now suppose that the response variable scores are transformed according to the formula
$\mathrm{y}^{\star\star}=5(\mathrm{y}+2)$ and that $\mathrm{y}^{\star\star}$ is regressed on $\mathrm{x}$.  Without redoing the regression calculation in detail, find $SE(\hat{\beta}^{\star\star})$ and $t^{\star\star}_0= \hat{\beta}^{\star\star}/SE(\hat{\beta}^{\star\star})$.


6. In general, how are the hypothesis tests and confidence intervals for $\beta$ affected by linear transformations of $\mathrm{y}$ and $\mathrm{x}$?



		
# Feedback comments etc.

If you have any comments about the homework, or the class, please write your feedback here.  We love to hear your opinions.

