---
title: "homework 07"
author: "Luyin Fu"
date: "November 3, 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,out.width="0.9\\linewidth",dev="png",fig.align  = 'center',warning = F,message = F,dpi = 500)
pacman::p_load(ggplot2,knitr,arm,data.table,foreign,gridExtra,car,stringr,rstan,rstanarm,zoo,magrittr,dplyr,tidyr,kableExtra)
```



# Data analysis 


## CD4 percentages for HIV infected kids

The folder `cd4` has CD4 percentages for a set of young children with HIV who were measured several times over a period of two years. The dataset also includes the ages of the children at each measurement.

```{r,echo=FALSE}
# Read in the data from an excel-format ".csv" file
hiv.data.raw <- fread ("http://www.stat.columbia.edu/~gelman/arm/examples/cd4/allvar.csv")

invisible(hiv.data.raw[,ok := !is.na(CD4PCT) ])

hiv.data<-hiv.data.raw[ok==TRUE]
invisible(hiv.data[,y :=sqrt (CD4PCT)])
 # kid's age (yrs) at the beginning of the study
invisible(hiv.data[,age.baseline := baseage ]  ) 
# kids age (yrs) at the time of measurement
invisible(hiv.data[,age.measurement := visage ] ) 
invisible(hiv.data[,time := visage - baseage ] )
setnames(hiv.data,"treatmnt","treatment") 
hiv.data<-hiv.data[complete.cases(hiv.data[,list(y,time,age.baseline,treatment)])]
```

1. Graph the outcome (the CD4 percentage, on the square root scale) for each child as a function of time.
```{r}
ggplot(hiv.data)+
  geom_line(aes(x=time, y=sqrt(CD4PCT),group = newpid))
```

2. Each child's data has a time course that can be summarized by a linear fit. Estimate these lines and plot them for all the children.

```{r}
fit1 = lm(formula=y~time+factor(newpid)-1, data = hiv.data)
ggplot(hiv.data)+
  geom_point(aes(x=time, y=y, color = factor(newpid), alpha=0.8))+
  geom_smooth(aes(x=time, y=y,color = factor(newpid)),method = "lm",se = F)+
  theme(legend.position = "none")
```


3. Set up a model for the children's slopes and intercepts as a function of the treatment and age at baseline. Estimate this model using the two-step procedure–first estimate the intercept and slope separately for each child, then fit the between-child models using the point estimates from the first step.
```{r}
intm <- hiv.data %>%
  dplyr::select(newpid, age.baseline, treatment) %>%
  distinct()
fit1.coef <- data.frame(intm,fit1$coefficients[2:length(fit1$coefficients)]) 
colnames(fit1.coef) <- c("newpid", "age.baseline", "treatment","coef.id")
rownames(fit1.coef) <- 1:250
fit2 = lm(data = fit1.coef,coef.id~age.baseline+factor(treatment)) 
summary(fit2)
```


4. Write a model predicting CD4 percentage as a function of time with varying intercepts across children. Fit using `lmer()` and interpret the coefficient for time.
```{r}
fit3 = lmer (y ~ time + (1|newpid), data = hiv.data) 
summary(fit3)
```

The coeffcient for time is the same for all children. With every unit increase in time, the square root of CO4 percentage is expected to reduce by 0.3661


5. Extend the model in (4) to include child-level predictors (that is, group-level predictors) for treatment and age at baseline. Fit using `lmer()` and interpret the coefficients on time, treatment, and age at baseline.

```{r}
fit4 <- lmer (y ~ time + factor(treatment) + age.baseline + (1 | newpid), data = hiv.data) 
summary(fit4)
```

\qquad Coefficient for time: with every unit increase in time, the square root of CO4 percentage is expected to reduce by 0.135 for children with the same treatment, baseline age.  

\qquad Coefficient for treatment: the average difference of CO4 percentage square root is 0.18 for childern in treatment 2 and treatment 1, if they have the same baseline age and at the same time point.  

\qquad Coefficient for baseline age: with every unit increase in aseline age, the square root of CO4 percentage is expected to reduce by 0.727 for children with the same treatment and at the same time point.  


6. Investigate the change in partial pooling from (4) to (5) both graphically and numerically.

```{r}
co.fit.3 = data.frame(fixef(fit3))
co.fit.3 = co.fit.3 %>% mutate(coef = rownames(co.fit.3))
co.fit.4 = data.frame(fixef(fit4))
co.fit.4 = co.fit.4 %>% mutate(coef = rownames(co.fit.4))
co.fit = full_join(co.fit.3,co.fit.4,by = "coef")
rownames(co.fit) = co.fit$coef
co.fit = co.fit %>% select(-coef)
colnames(co.fit) = c("Model in (4)","Model in (5)")
kable(co.fit,align = "c",format = "latex",booktabs = T)
```

7. Use the model fit from (5) to generate simulation of predicted CD4 percentages for each child in the dataset at a hypothetical next time point.

```{r}
new.data = hiv.data %>% select(newpid,treatment,age.baseline) %>% distinct()
new.data = new.data %>% mutate(time = 2)
pre = predict(fit4,newdata = new.data,type = "response")
pre = cbind(pre,new.data)
colnames(pre) = c("y","newpid","treatment","age.baseline","time")
kable(head(pre,10),align = "c",format = "latex",booktabs = T)
```

8. Use the same model fit to generate simulations of CD4 percentages at each of the time periods for a new child who was 4 years old at baseline.

```{r}
time_point=sort(unique(hiv.data$time)) %>% data.frame()
colnames(time_point) = "time"
new.data = time_point %>% mutate(treatment = 2, age.baseline = 4)
pre = data.frame(predict(fit4,newdata = new.data,re.form = NA))
colnames(pre) = "y"
pre = cbind(pre,new.data)
kable(head(pre,10),align = "c",format = "latex",booktabs = T)
```

9. Posterior predictive checking: continuing the previous exercise, use the fitted model from (5) to simulate a new dataset of CD4 percentages (with the same sample size and ages of the original dataset) for the final time point of the study, and record the average CD4 percentage in this sample. Repeat this process 1000 times and compare the simulated distribution to the observed CD4 percentage at the final time point for the actual data.
```{r}
# new.data = hiv.data %>% select(newpid,treatment,age.baseline) %>% distinct()
# new.data = new.data %>% mutate(time = 1.9383333)
# pre = predict(fit4,newdata = new.data)
# new.data[paste0("pre",1)] = pre
# iter = hiv.data %>% select(newpid,treatment,age.baseline) %>% distinct()
# iter = iter %>% mutate(time = 1.9383333)
# pre.result = iter
# for(i in 1:1000){
#   pre = predict(fit4,newdata = iter)
#   pre.result[paste0("pre",i)] = pre
# }
#the final time measurement is different for each child
```

10. Extend the model to allow for varying slopes for the time predictor.
```{r}
#time is not a group level predictor
fit6 <- lmer(formula = y ~ factor(treatment) + age.baseline + time + (1 + time | newpid), 
             data = hiv.data)
summary(fit6)
```


11. Next fit a model that does not allow for varying slopes but does allow for different coefficients for each time point (rather than fitting the linear trend).
```{r}
fit7 <- lmer(formula = y ~ factor(treatment) + age.baseline + time + 
               (time - 1| newpid), 
             data = hiv.data)
```


12. Compare the results of these models both numerically and graphically.
```{r}
summary(fit6)
summary(fit7)

E=data.frame(coef(fit6)$newpid) 
E %<>% mutate(newpid=rownames(E))
ggplot(hiv.data)+
  geom_point(aes(x=time, y=y, color = factor(newpid), alpha=0.8))+
  geom_abline(intercept = E$X.Intercept., slope=E$time, color=factor(E$newpid), alpha=0.5)+
  theme(legend.position = "none")+
  ggtitle("varying slopes")

G=data.frame(coef(fit7)$newpid) 
G %<>% mutate(newpid=rownames(G))
ggplot(hiv.data)+
  geom_point(aes(x=time, y=y, color = factor(newpid), alpha=0.8))+
  geom_abline(intercept = G$X.Intercept., slope=G$time, color=factor(G$newpid), alpha=0.5)+
  theme(legend.position = "none")+
  ggtitle("different coefficients for each time point")
```


## Figure skate in the 1932 Winter Olympics

The folder olympics has seven judges' ratings of seven figure skaters (on two criteria: "technical merit" and "artistic impression") from the 1932 Winter Olympics. Take a look at 
http://www.stat.columbia.edu/~gelman/arm/examples/olympics/olympics1932.txt
```{r,echo=FALSE}
filename<- "http://www.stat.columbia.edu/~gelman/arm/examples/olympics/olympics1932.txt"
olympics1932_na<-read.fwf(filename,widths=c(2,14,9,9,9,9,9,9,9),skip=21,header = FALSE)
colnames(olympics1932_na)<- c("pair", "criterion", "judge_1",  "judge_2",  "judge_3",
                              "judge_4",  "judge_5" , "judge_6",  "judge_7")

olympics1932<-na.locf(olympics1932_na)
olympics1932$criterion<-str_trim(olympics1932_na$criterion)
```

1. Construct a $7\times 7 \times 2$ array of the data (ordered by skater, judge, and judging criterion).

```{r}
H1 <- olympics1932 %>% filter(criterion=="Program")
H2 <- olympics1932 %>% filter(criterion=="Performance")
Olympics1=array(dim = c(7,7,2))
Olympics1[,,1]=as.matrix(H1[-c(1,2)])
Olympics1[,,2]=as.matrix(H2[-c(1,2)])
print(Olympics1)
```

2. Reformulate the data as a $98\times 4$ array (similar to the top table in Figure 11.7), where the first two columns are the technical merit and artistic impression scores, the third column is a skater ID, and the fourth column is a judge ID.

```{r}
Olympics2 <- olympics1932 %>% 
  pivot_longer(-c(1,2), names_to = "judge ID", values_to="score") %>%
  pivot_wider(names_from = "criterion", values_from="score") %>%
  separate(col = `judge ID`, into = c("nevermind", "judge"), sep = "_", remove = T) %>%
  select(Program, Performance, pair, judge)
kable(head(Olympics2,10), align = "c",format = "latex",booktabs = T)

```

3. Add another column to this matrix representing an indicator variable that equals 1 if the skater and judge are from the same country, or 0 otherwise.

```{r, warning=F}
data = Olympics2
data = data %>% mutate(same.country = ifelse(data$pair==as.numeric(data$judge),1,0))
```

4. Write the notation for a non-nested multilevel model (varying across skaters and judges) for the technical merit ratings and fit using lmer().
```{r}
fit8 <- lmer(Program ~  (1 | pair) + (1 | judge), data = data)
summary(fit8)
```

5. Fit the model in (4) using the artistic impression ratings.
```{r}
fit9 <- lmer(Performance ~ (1 | pair) + (1 | judge) ,data = data)
summary(fit9)
```

6. Display your results for both outcomes graphically.

```{r}
H=data.frame(coef(fit8)$judge) 
H %<>% mutate(pair=rownames(H))
ggplot(data)+
  geom_point(aes(x=as.factor(pair), y=Program, color=as.factor(pair)))+
  geom_abline(intercept = H$X.Intercept., slope= rep(0,7), color=H$pair, alpha=0.5)+
  scale_x_discrete(breaks = seq(1,7,1))

I=data.frame(coef(fit9)$judge) 
I %<>% mutate(pair=rownames(I))
ggplot(data)+
  geom_point(aes(x=as.factor(pair), y=Performance, color=as.factor(pair)))+
  geom_abline(intercept = I$X.Intercept., slope= rep(0,7), color=I$pair, alpha=0.5)+
  scale_x_discrete(breaks = seq(1,7,1))
```

7. (optional) Use posterior predictive checks to investigate model fit in (4) and (5).

```{r}

```

## Different ways to write the model:

Using any data that are appropriate for a multilevel model, write the model in the five ways discussed in Section 12.5 of Gelman and Hill.  

\qquad Use the Olympic data as example and we set the technical merit as response variable, and with no predictors, treat each athelete as a group with score from each judge, the score of each athelete is $y_{ij}$ means the score of athelete i from judge j.  

+ model in way 1: $y_{ij} = N(\alpha_i,\sigma^2), \alpha_i = \rho + \rho_1u_i+\varepsilon, \varepsilon = N(0,\sigma_{\alpha}^2)$, here the $\alpha_i$ is the mean of each athlete, the score of each athlete comply with normal distribution, at the same time, the mean of each athlete also comply with normal distribution, but the mean are different cross athletes;

+ model in way 2: $y_{ij} = N(\alpha_i,\sigma^2), \alpha_i = N(\rho + \rho_1u_i,\sigma_{\alpha}^2)$, this model is exactly the same with way 1, just in way 1, the variability of $\alpha_i$ is introduced by $\varepsilon_i$, here we directly represent $\alpha_i$ as a random variable.  

## Models for adjusting individual ratings: 

A committee of 10 persons is evaluating 100 job applications. Each person on the committee reads 30 applications (structured so that each application is read by three people) and gives each a numerical rating between 1 and 10.


1. It would be natural to rate the applications based on their combined scores; however, there is a worry that different raters use different standards, and we would like to correct for this. Set up a model for the ratings (with parameters for the applicants and the raters).


2. It is possible that some persons on the committee show more variation than others in their ratings. Expand your model to allow for this.


